Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-22 WSSSPE1_report/WSSSPE1_summary_report.tex
review and small changes to sec 9

\subsection{Scientific software}
Scientific software consists primarily of special-purpose software that was
developed for a target use-case scenario or fixed/frozen requirements in mind.
Software projects pertaining to specific scientific domain often tend to be in
a niche and the user community tends to be small to medium. Specific
requirements such as numerical accuracy and algorithmic optimization are some
of the paramount requirements of most scientific software. The resulting
codebase is relatiely small. A smaller codebase and fixed requirements result
in stability, ease of installation, and configurability. Many such projects are
treated as libraries used into larger systems such as some of the utility
software discussed in the previous section. While the software can stay stable
and require relatively low maintenance, the responsibility is often on the
shoulders of a very few developers. Community participation in the development
tends to be low. Development is linear and simplistic with a limited scope to
follow software best practices. Domain science oriented software is developed
to solve  specific problems, e.g., \emph{DUNE}, \emph{PETSc},
\emph{MINRES-QLP}, \emph{FASTMath}. Sustainability of such software is often a
significant challenge. Many submissions reported the software is considered a
`byproduct' of the actual research.  Others contended that the software was not
the main funded part of their research. One exception was \emph{Kitware}, which
while being a software product specializing in the scientific process, has a
core focus of developing communities around software processes. One example of
this process is the development of the \emph{CMake} build utility, which
started out as a building tool for \emph{ITK} but grew to become a generic
build utility for C++ projects.

>>>>

\subsection{Scientific software}
Scientific software consists primarily of special-purpose software that was
developed for a target use-case scenario or fixed/frozen requirements in mind.
Software projects pertaining to specific scientific domain often tend to be in
a niche and the user community tends to be small to medium.
%
Specific needs such as numerical accuracy and algorithmic optimization are some
of the paramount requirements of most scientific software.
%
The resulting codebase is relatively small. A smaller codebase and fixed
requirements result in stability, ease of installation, and configurability.
Many such projects are treated as libraries to be used into larger systems such as
some of the utility software discussed in the previous section.
%
While the
software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a very few developers.
%
Community
participation in the development tends to be low. Development is linear and
simplistic with a limited scope to follow software best practices.
%
Some examples of such software discussed as part of the workshop are,
\emph{DUNE}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}.
%
Sustainability of such software is often a significant challenge. Many
submissions reported the software is considered a `byproduct' of the actual
research. Others contended that the software was not the main funded part of
their research. One exception was \emph{Kitware}, which while being a software
product specializing in the scientific process, has a core focus of developing
communities around software processes. An instance of this process is the
development of the \emph{CMake} build utility, which started out as a building
tool for \emph{ITK} but grew to become a generic build utility for C++
projects.
